Outline paper - Multiple imputation for statistical disclosure control: Creating synthetic datasets with *mice*

### 2021-04-08 mice_synthesizing/paper.Rmd
Small adjustment to intro, and some additions to regular body of text.

Open science, including open data, has been marked as the future of science [@gewin_data_2016], and the advantages of publicly available research data are numerous [@molloy_open_2011; @walport_brest_sharing_2011]. For example, collecting research data requires an enormous investment both in terms of time and monetary resources. Openly accessible research data bears the potential of increasing the scientific returns for the same data collection effort. Additionally, the fact that public funds are used for data collection results in increasing demand for the collected data. Nevertheless, the possibilities to share research data directly are often very limited due to restrictions with regard to privacy and confidentiality.

A quick approach to lessen the identities in the data is anonimization. However, simply anonimizing the data is not sufficient to fulfil contemporary privacy and confidentiality  requirements [@ohm_broken_2009]. Over the years, several other techniques have been used to increase the confidentiality of the data, such as categorizing continuous variables, top coding values above a certain threshold or adding random noise to the observed values [@drechsler_synthetic_2011]. However, these methods may distort relationships between variables, thereby reducing the data quality and the scientific returns for re-using the same data for further research.

@@

@@

This approach can be straightforwardly extended to a synthetic data framework. However, rather than imputing missing data, the actually observed data is overimputed, that is, the observed values are replaced by multiple imputations. For simplicity, assume that the data is completely observed (i.e., $\textbf{Y} = \textbf{Y}_{obs}$). Following the notation of @reiter_raghunathan_multiple_2007, let $Z_j = 1$ if any of the values of unit $j$ are to be replaced by imputations, and $Z_j = 0$ otherwise, with $\textbf{Z} = (Z_1, Z_2, \dots, Z_n)$. Accordingly, the data consists of values that are to be replaced and values that are to be kept (i.e., $\textbf{Y}^{(l)} = (\textbf{Y}^{(l)}_{rep}, \textbf{Y}_{nrep})$. Note the superscript $^{(l)}$ indicating the $l$th synthetic data set, with $l = 1, 2, \dots, m$. Now, instead of imputing $\textbf{Y}^{(l)}_{mis}$ with draws from the posterior predictive distribution of $P(Y_{j, mis} | \textbf{Y}_{obs}, \theta)$ as in the missing data case, $\textbf{Y}^{(l)}_{rep}$ is imputed from the posterior distribution of $(Y_{rep} | \textbf{Y}, \textbf{Z}, \theta)$. This process results in the synthetic data $\textbf{D} = (\textbf{D}^{(1)}, \textbf{D}^{(2)}, \dots, \textbf{D}^{(m)})$.

By default, `mice` imputes all values that are coded as missing in the dataset. However, using the `where` parameter within the `mice`-function, one can choose to impute any value in the dataset. As input, the `where` parameter requires a matrix of the same dimensions as the data, (i.e., a $n \times k$ matrix, with $n$ the number of observations in the collected data) containing logicals $z_{jk}$ that indicate which cells are selected to have their values replaced. This approach allows to *overimpute* a subset of the observed data, or the observed data as a whole, resulting in a dataset that partly or completely consists of synthetic values. Other parameters that are required to be specified are the imputation method (`method`), the number of imputations ($m$) and the predictor matrix, specifying the predictor variables that are used in the imputation model for every variable $j$ that has some of its values replaced (`predictorMatrix`).

@@

Additional parameters that come with the use of CART are the complexity parameter `cp` and the minimum number of observations in any terminal node `minbucket`, that both restrain the flexibility of the imputation model. Appropriate values for these parameters, as well as the input for the `predictorMatrix`, depend on the data at hand. Hence, the specification of these parameters will be illustrated using an actual dataset. The `boys` data [@fredriks_boys_2000] that is in-built within the `mice` package is used to illustrate the specification of these parameters accordingly. **DETERMINE WHETHER THIS DATASET SUFFICES; BECAUSE IMPUTATION MIGHT MAKE THE DATA ARTIFICIALLY SIMILAR BEFORE SYNTHESIZING, MAKING SYNTHESIZING TOO EASY** Unfortunately, this dataset does not differ from the vast majority of collected datasets, in the sense that it suffers from missing data as well. For simplicity, the data is completed using the default `mice` imputation model to create a single, fully observed dataset. The data consists of the values of $`r nrow(mice::boys)`$ Dutch boys on the following $`r ncol(mice::boys)`$ variables: age (in years), height (in cm), weight (in kg), body mass index, heid circumference (in cm), Genital Tanner stage (G1-G5), testicular volume (in ml) and region.

The values of the parameters `cp` and `minbucket` ought to adhere to the call for imputation models that are as flexible as possible. Accordingly, the complexity parameter is fixed at `cp = 1e-08` rather than the default value `1e-08` and the minimum number of observations in each terminal node is set at `minbucket` $= 3$ rather than the default value $5$. The bias and variance of the final estimates justified these decisions. The predictor matrix was specified in such a way that 
Additionally, the number of imputations was set to $m = 5$.

>>>>

Open science, including open data, has been marked as the future of science [@gewin_data_2016], and the advantages of publicly available research data are numerous [@molloy_open_2011; @walport_brest_sharing_2011]. Collecting research data requires an enormous investment both in terms of time and monetary resources. Openly accessible research data bears the potential of increasing the scientific returns for the same data collection effort. Additionally, the fact that public funds are used for data collection results in increasing demand for the collected data. Nevertheless, the possibilities to distribute research data directly are often very limited due to restrictions with regard to privacy and confidentiality. Altough these regulations are much needed, privacy constraints are also ranked among the toughest challenges to overcome in the advancement of computational social science [@lazer_life_2009].

Anonimizing the data might seem a quick and appealing approach to limit the identifyability of participants. However, this approach is not sufficient to fulfil contemporary privacy and confidentiality requirements [@ohm_broken_2009; @national_putting_2007]. Over the years, several other techniques have been used to increase the confidentiality of the data, such as categorizing continuous variables, top coding values above a certain threshold or adding random noise to the observed values [@drechsler_synthetic_2011]. However, these methods may distort relationships between variables, thereby reducing the data quality and the scientific returns for re-using the same data for further research.

@@

**NEXT TWO PARAGRAPHS ARE PROBABLY NOT CORRECT YET; NOTE TO SELF: UPDATE THIS**

@@

This approach can be straightforwardly extended to a synthetic data framework. However, rather than imputing missing data, the actually observed data is overimputed, that is, the observed values are replaced by multiple imputations. For simplicity, assume that the data is completely observed (i.e., $\textbf{Y} = \textbf{Y}_{obs}$). Following the notation of @reiter_raghunathan_multiple_2007, let $Z_i = 1$ if any of the values of unit $i$ are to be replaced by imputations, and $Z_i = 0$ otherwise, with $\textbf{Z} = (Z_1, Z_2, \dots, Z_n)$. Accordingly, the data consists of values that are to be replaced and values that are to be kept (i.e., $\textbf{Y} = (\textbf{Y}_{rep}, \textbf{Y}_{nrep})$. Now, instead of imputing $\textbf{Y}_{mis}$ with draws from the posterior predictive distribution of $P(Y_{j, mis} | \textbf{Y}_{obs}, \theta)$ as in the missing data case, $\textbf{Y}_{rep}$ is imputed from the posterior distribution of $P(Y^{(l)}_{j, rep} | \textbf{Y}^{(l)}_{-j}, \textbf{Z}, \theta)$, where $l$ is an indicator for the synthetic dataset ($l = 1, 2, \dots, m$). Note that synthetic values that are imputed at an earlier step can be used for imputing variable $j$. This process results in the synthetic data $\textbf{D} = (\textbf{D}^{(1)}, \textbf{D}^{(2)}, \dots, \textbf{D}^{(m)})$.

By default, `mice` imputes all values that are coded as missing in the dataset. However, using the `where` parameter within the `mice`-function, one can choose to impute any value in the dataset. As input, the `where` parameter requires a matrix of the same dimensions as the data, (i.e., a $n \times k$ matrix, with $n$ the number of observations in the collected data) containing logicals $z_{ij}$ that indicate which cells are selected to have their values replaced. This approach allows to *overimpute* a subset of the observed data, or the observed data as a whole, resulting in a dataset that partly or completely consists of synthetic values. Other parameters that are required to be specified are the imputation method (`method`), the number of imputations ($m$) and the predictor matrix, specifying the predictor variables that are used in the imputation model for every variable $j$ that has some of its values replaced (`predictorMatrix`).

@@

Additional parameters that come with the use of CART are the complexity parameter `cp` and the minimum number of observations in any terminal node `minbucket`, that both restrain the flexibility of the imputation model. Appropriate values for these parameters, as well as the input for the `predictorMatrix`, depend on the data at hand. Hence, the specification of these parameters will be illustrated using an actual dataset. The `boys` data [@fredriks_boys_2000] that is in-built within the `mice` package is used to illustrate the specification of these parameters accordingly. **DETERMINE WHETHER THIS DATASET SUFFICES; BECAUSE IMPUTATION MIGHT MAKE THE DATA ARTIFICIALLY SIMILAR BEFORE SYNTHESIZING, MAKING SYNTHESIZING TOO EASY** Unfortunately, this dataset does not differ from the vast majority of collected datasets, in the sense that it suffers from missing data. For simplicity, the data is completed using the default `mice` imputation model to create a single, fully observed dataset. The data consists of the values of $`r nrow(mice::boys)`$ Dutch boys on the following $`r ncol(mice::boys)`$ variables: age (in years), height (in cm), weight (in kg), body mass index (BMI), head circumference (in cm), Genital Tanner stage (G1-G5), testicular volume (in ml) and region.

The values of the parameters `cp` and `minbucket` ought to adhere to the call for imputation models that are as flexible as possible. Accordingly, the complexity parameter is fixed at `cp = 1e-08` rather than the default value `1e-08` and the minimum number of observations in each terminal node is set at `minbucket` $= 3$ rather than the default value $5$. By allowing for more complexity in the imputation model, bias in the estimates from the synthetic dataset is reduced. The predictor matrix was specified so that BMI was not imputed from a model, but calculated from the imputed height and weight values ($BMI = kg / m^2$), an approach called passive imputation. Additionally, BMI is excluded as a predictor for weight and height, because BMI values are directly dependent on the values of weight and height. All other variables $Y_{-j}$ are used as predictors for variable $Y_j$. Additionally, the number of imputations is set to $m = 5$.
